---
title: "Indiana Feeding Rate Assays"
author: "Meghan Duffy"
date: "October 27, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", error = TRUE)
```

# Does feeding rate differ before and after epidemics?

Data collection for this project led by Katie Hunsberger (technician in Duffy Lab), using clones sent from Spencer Hall's lab in Indiana. Data analysis led by Meghan Duffy (duffymeg@umich.edu)

```{r,load libraries,message=FALSE,echo=FALSE}
library(ggplot2)
library(dplyr)
library(cowplot)
library(stringr)
```
Right now, data sheet just has Island data in it. Will need to update this with rest of data, but starting with this as a trial.

Note about data: The fluorometer drifts over time. We quantify this by running controls periodically. We then need to factor out this drift to determine how the baseline RFUs for an ungrazed tube shifts over time. 

## Jenny Bryan's (elegant!) proposal for dealing with fluorometer drift issue

Read data in.

```{r, load data}
df <- read.csv("indianafeedingrateassays_withcontrols.csv", na.strings = ".")
df$Block <- factor(df$Block)
head(df)
str(df)
```

For the controls only, plot RFU against order and color by Block. Add a fitted line for each Block.

```{r, plot controls}
df %>%
  filter(CloneSubline == "Control") %>% 
  ggplot(aes(x = order, y = RFU, color = Block)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() 
```

Regress RFU on order within Block, for the controls. Eyeball-o-metrically check the intercept and slope against the plot.

```{r, regression of controls}
control_fit <- lm(RFU ~ Block/order - 1, data = df,
                  subset = CloneSubline == "Control")
summary(control_fit)
```

Add a variable into `df` for the predicted RFU of a control tube; this tells us what RFUs we'd expect if there was no Daphnia in a tube (or if the Daphnia didn't eat anything)

```{r, calculate predicted control RFUs}
df$predictedControlRFU <-
  predict(control_fit, newdata = df[c("order", "Block")])
head(df)
```

Now use that to calculate the clearance rate (which we also call feeding rate, because "clearance" has a different meaning to disease ecologists, and so can be a confusing term). Also need to know that the assays were run for 3 hours (0.125 days) and in 15 mL (0.015 L) water.

```{r, calculate clearance rate}
df <-
  mutate(df, CR = log(predictedControlRFU/RFU)*0.015*0.125)
head(df)
```

Now calculate the size-specific clearance rate
```{r, calculate size-specific clearance rate}
df <-
  mutate(df, SSCR = CR/(length_mm)^2)
head(df)
```

Check to make sure there don't seem to be major trends in the clearance rates over the course of the assay
```{r, plot clearance rate}
df %>%
  ggplot(aes(x = order, y = CR, color = Block)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() 
```

Hrm, something looks not right with block 1. Will need to look into that more. First check: is something going wrong with the predictions? Testing this by subsetting the data so there's just one block.

```{r, subset to just block 1}
testdf <- read.csv("indianafeedingrateassays_withcontrols.csv", na.strings = ".")
testdf$Block <- factor(testdf$Block)
testdf <- subset(testdf, Block == 1)
tail(testdf)
```

```{r, regression of block 1 controls}
control_fit2 <- lm(RFU ~ order, data = testdf,
                  subset = CloneSubline == "Control")
summary(control_fit2)
```

```{r, calculate predicted control RFUs for just block 1}
testdf$predictedControlRFU <-
  predict(control_fit2, newdata = testdf[c("order", "Block")])
head(testdf)
```

```{r, calculate clearance rate for just block 1}
testdf <-
  mutate(testdf, CR = log(predictedControlRFU/RFU)*0.015*0.125)
head(testdf)
```

```{r, plot clearance rate for just block 1}
testdf %>%
  ggplot(aes(x = order, y = CR)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() 
```

Okay. It's not just a problem with calculating predicted RFUs. Will need to figure out whether we should do anything about this later. The first step might be seeing if it happened in the blocks for other lakes, too.

##Looking at data for different genotypes and sublines

First step: separate clone and subline (which are currently merged into the "CloneSubline" variable)

```{r, separate clone and subline}
dfnc <- subset(df,CloneSubline != "Control")
dfnc$CloneSubline <- as.character(dfnc$CloneSubline)
dfnc$Subline <- str_sub(dfnc$CloneSubline,-1,length(dfnc$CloneSubline))
dfnc$Clone <- substr(dfnc$CloneSubline, 1, nchar(dfnc$CloneSubline)-1)
dfnc$Subline
dfnc$Clone
```

Then, figure out the mean among the different replicates for each clone*subline combination
Seeing if I can figure out how to graph it first

```{r, plotting data for different clone*subline combinations}
ggplot(dfnc,aes(x=Clone,y=SSCR)) +
  geom_point(aes(color=prepost,shape=Subline))
```

Calculating one value for each clone; first, calculating a mean for each clone-subline combination, then averaging those two values to get one for each clone.
```{r, mean for each clone}
meanbysublineSSCR <- dfnc %>%
  subset(!is.na(SSCR)) %>%
  group_by(Clone,Subline,prepost) %>%
  summarise(avg=mean(SSCR))

meanbycloneSSCR <- meanbysublineSSCR %>%
  group_by(Clone,prepost) %>%
  summarise(avg=mean(avg))

meanbysublineCR <- dfnc %>%
  subset(!is.na(CR)) %>%
  group_by(Clone,Subline,prepost) %>%
  summarise(avg=mean(CR))

meanbycloneCR <- meanbysublineCR %>%
  group_by(Clone,prepost) %>%
  summarise(avg=mean(avg))

meanbysublinelength <- dfnc %>%
  subset(!is.na(length_mm)) %>%
  group_by(Clone,Subline,prepost) %>%
  summarise(avg=mean(length_mm))

meanbyclonelength <- meanbysublinelength %>%
  group_by(Clone,prepost) %>%
  summarise(avg=mean(avg))
```

Plotting the means
```{r, plots of mean SSCR and length for different clones}
ggplot(meanbycloneSSCR,aes(x=prepost,y=avg)) +
  geom_point() +
  ylab("Size-specific clearance rate") +
  ggtitle("Mean Size-specific Clearance Rate; each point is one genotype") +
  scale_x_discrete(limits=c("pre", "post", "std")) +
  xlab("Pre- vs. post-epidemic")

ggplot(meanbycloneSSCR,aes(x=prepost,y=avg)) +
  geom_boxplot() +
  ylab("Size-specific clearance rate") +
  ggtitle("Mean Size-specific Clearance Rate (box plot shows averages across genotypes)") +
  scale_x_discrete(limits=c("pre", "post", "std")) +
  xlab("Pre- vs. post-epidemic")

ggplot(meanbycloneCR,aes(x=prepost,y=avg)) +
  geom_point() +
  ylab("Clearance rate") +
  ggtitle("Mean Clearance Rate; each point is one genotype") +
  scale_x_discrete(limits=c("pre", "post", "std")) +
  xlab("Pre- vs. post-epidemic")

ggplot(meanbycloneCR,aes(x=prepost,y=avg)) +
  geom_boxplot() +
  ylab("Clearance rate") +
  ggtitle("Mean Clearance Rate (box plot shows averages across genotypes)") +
  scale_x_discrete(limits=c("pre", "post", "std")) +
  xlab("Pre- vs. post-epidemic")

ggplot(meanbyclonelength,aes(x=prepost,y=avg)) +
  geom_point() +
  ylab("Length (mm)") +
  ggtitle("Mean length; each point is one genotype") +
  scale_x_discrete(limits=c("pre", "post", "std")) +
  xlab("Pre- vs. post-epidemic")

ggplot(meanbyclonelength,aes(x=prepost,y=avg)) +
  geom_boxplot() +
  ylab("Length (mm)") +
  ggtitle("Mean length (box plot shows averages across genotypes)") +
  scale_x_discrete(limits=c("pre", "post", "std")) +
  xlab("Pre- vs. post-epidemic")
```

To do next: analyze evolution, looking for changes in mean and/or variance
